A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-07-29 survey.tex
proof-read section III

Systems that play RTS games need to address most, if not all, the aforementioned problems together. Therefore, it is hard to classify existing work on RTS AI as addressing the different problems above. For that reason, in order to classify existing work on RTS AI, we will divide it according to three levels of abstraction: strategy (which loosely corresponds to ``macro''), tactics and reactive control (which loosely corresponds to ``micro'').

@@

% HARD-CODED:
Hard-coded approaches have been extensively used in commercial RTS games. The most common approaches use finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003} in order to let the AI author hard-code the strategy that the AI will employ. The idea behind FSMs is to decompose the AI behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Commercial approaches also include Hierarchical FSMs, in which FSMs are composed hierarchically. These hard-coded approaches have achieved a significant amount of success, and, as we will discuss later, have also been used in many academic RTS AI research systems, as discussed in Section \ref{sec:bot}. However, these hard-coded approaches struggle to encode dynamic, adaptive behaviors, and are easily exploitable by adaptive opponents.

@@

% PLANNING:
Approaches using planning techniques have also been explored in the literature. For example Onta\~{n}\'{o}n et al. \cite{CBR_Planning} explored the use of real-time case-based planning (CBP) in the domain of Wargus (a Warcraft II clone). In their work, they used human demonstration to learn plans, which are then composed at run-time in order to form full-fledges strategies to play the game. In \cite{PlanRetrieval} they improve over  their previous CBP approach by using situation assessment for improving the quality and speed of plan retrieval. Hierarchical Task-Network (HTN) planning has also been explored with some success in the context of simpler first-person shooter games \cite{HTNPlanning}. Planning approaches offer more adaptivity of the AI strategy compared to hard-coded approaches. However, the real-time constraints of RTS games limit the planning approaches that can be applied, HTN and case-based planning being the only ones explored so far. Moreover, none of these approaches addresses any timing or scheduling issues, which are key in RTS games. On notable exception is the work of Churchill and Buro \cite{churchill2011build}, who used planning in order to construct its economic build-orders taking into account timing constraints of the different actions.

% MACHINE LEARNING:
Concerning machine learning-based approaches, Weber and Mateas \cite{WeberCig09} proposed a data mining approach to strategy prediction and performed supervised learning on labeled StarCraft replays. Dereszynski et al. \cite{HMMstrat_RTS_AIIDE11} used Hidden Markov Models (HMM) to learn the transition probabilities of sequences of building construction orders and kept the most probable ones to produce probabilistic behavior models (in StarCraft). Synnaeve and Bessi\`{e}re \cite{SynnaeveOpeningCig11} used the dataset of \cite{WeberCig09} and presented a Bayesian semi-supervised model to learn from replays and predict openings (early game strategies) from StarCraft replays. The openings are labeled by EM clustering considering appropriate features. Then, in \cite{SynnaeveAIIDE11}, they presented an unsupervised learning Bayesian model for tech-tree prediction, still using replays. % Finally \cite{GemineImitative} used the same ``build-order features'' approach to learn to adapt the bot's build order to the opponents strategy in StarCraft 2.
Finally, evolutionary approaches to determine priorities of high level tasks was explored by Young and Hawes in their QUORUM system \cite{young2012evolutionary}, showing improvement over static priorities.

@@

% MACHINE LEARNING:
Concerning tactical decision making, many different approaches have been explored such as machine learning or game tree search. 
{\color{blue} Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters for unit tracking, although the used first-person shooter (FPS) games for their experimentation, the results apply to RTS games as well}. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. \cite{SynnaeveTactics} combined space abstraction into regions from \cite{Perkins10} and tactical-decision making by assigning scores (economical, defenses, etc.) to regions and looking for their correspondences to tactical moves (attacks) in pro-gamers replays.
{\color{blue} Finally, Christopher Miles \cite{miles2006co} created the idea of {\em IMTrees}, a tree where each leaf node is an influence map, and each intermediate node is a combination operation (sum, multiplication); Miles used evolutionary algorithms to learn IMTrees for each strategic decision in the game involving spatial reasoning by combining a set of basic influence maps.}

@@

Additionally, there has been some interesting uses of reinforcement learning (RL) \cite{Sutton}: 
Wender and Watson \cite{WenderRL} evaluated the different major RL algorithms for (decentralized) micro-management, which perform all equally. Marthi et al. \cite{Marthi05} employ concurrent hierarchical Q-learning (units Q-functions are combined at the group level) RL to efficiently control units in a ``one robot with multiple effectors'' fashion. Madeira et al. \cite{Madeira06} advocate the use of prior domain knowledge to allow faster RL learning and applied their work on a turn-based strategy game. This is because the action space to explore is gigantic for real game setups. It requires exploiting the existing structure of the game in a partial program (or a partial Markov decision process) and a shape function (or a heuristic) \cite{Marthi05}. Another approach is that proposed by Jaide and Mu{\~n}oz-Avila \cite{jaidee2012classq} through learning just one Q-function for each unit type, in order to cut down the search space. Other approaches that aim at learning the parameters of an underlying model have also been explored. For example Ponsen and Spronck \cite{GA} used evolutionary learning techniques, but face the same problem of dimensionality. For example, evolutionary optimization by simulating fights can easily adapted to any parameter-dependent micro-management control model, as shown by \cite{OthmanSimu} which optimizes an AIIDE 2010 micro-management competition bot.

@@

Holistic approaches to address RTS AI attempt to address the whole problem using a single unified method. To the best of our knowledge, with a few exceptions, such as the Darmok system \cite{OntanonMSR10} (which uses a combination of case-based reasoning and learning from demonstration) or ALisp \cite{Marthi05}, there hasn't been much work in this direction.   The main reason is that the complexity of RTS games is too large, and approaches that decompose the problem into smaller, separate, problems, achieve better results in practice. However, holistic approaches, based, for example, on Monte Carlo Tree Search, have only been explored in the context of smaller-scale RTS games~\cite{ontanon2013naive}, but techniques that scale up to large RTS games like StarCraft are still not known.

>>>>

Systems that play RTS games need to address most, if not all, the aforementioned problems together. Therefore, it is hard to classify existing work on RTS AI as addressing the different problems above. For that reason, %in order to classify existing work on RTS AI, 
we will divide it according to three levels of abstraction: strategy (which loosely corresponds to ``macro''), tactics and reactive control (which loosely corresponds to ``micro'').

@@

% HARD-CODED:
Hard-coded approaches have been extensively used in commercial RTS games. The most common ones use finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003} in order to let the AI author hard-code the strategy that the AI will employ. The idea behind FSMs is to decompose the AI behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Commercial approaches also include Hierarchical FSMs, in which FSMs are composed hierarchically. These hard-coded approaches have achieved a significant amount of success, and, as we will discuss later in Section~\ref{sec:bot}, have also been used in many academic RTS AI research systems. However, these hard-coded approaches struggle to encode dynamic, adaptive behaviors, and are easily exploitable by adaptive opponents.

@@

% PLANNING:
Approaches using planning techniques have also been explored in the literature. For example Onta\~{n}\'{o}n et al. \cite{CBR_Planning} explored the use of real-time case-based planning (CBP) in the domain of Wargus (a Warcraft II clone). In their work, they used human demonstration to learn plans, which are then composed at run-time in order to form full-fledges strategies to play the game. In \cite{PlanRetrieval} they improve over  their previous CBP approach by using situation assessment for improving the quality and speed of plan retrieval. Hierarchical Task-Network (HTN) planning has also been explored with some success in the context of simpler first-person shooter games \cite{HTNPlanning}. Planning approaches offer more adaptivity of the AI strategy compared to hard-coded approaches. However, the real-time constraints of RTS games limit the planning approaches that can be applied, HTN and case-based planning being the only ones explored so far. Moreover, none of these approaches addresses any timing or scheduling issues, which are key in RTS games. On notable exception is the work of Churchill and Buro \cite{churchill2011build}, who used planning in order to construct its economic build-orders, taking into account timing constraints of the different actions.

% MACHINE LEARNING:
Concerning machine learning-based approaches, Weber and Mateas \cite{WeberCig09} proposed a data mining approach to strategy prediction and performed supervised learning on labeled StarCraft replays. Dereszynski et al. \cite{HMMstrat_RTS_AIIDE11} used Hidden Markov Models (HMM) to learn the transition probabilities of sequences of building construction orders and kept the most probable ones to produce probabilistic behavior models (in StarCraft). Synnaeve and Bessi\`{e}re \cite{SynnaeveOpeningCig11} used the dataset of \cite{WeberCig09} and presented a Bayesian semi-supervised model to learn from replays and predict openings (early game strategies) from StarCraft replays. The openings are labeled by EM clustering considering appropriate features. Then, in \cite{SynnaeveAIIDE11}, they presented an unsupervised learning Bayesian model for tech-tree prediction, still using replays. % Finally \cite{GemineImitative} used the same ``build-order features'' approach to learn to adapt the bot's build order to the opponents strategy in StarCraft 2.
Finally, evolutionary approaches to determine priorities of high level tasks were explored by Young and Hawes in their QUORUM system \cite{young2012evolutionary}, showing improvement over static priorities.

@@

% MACHINE LEARNING:
Concerning tactical decision making, many different approaches have been explored such as machine learning or game tree search. 
{\color{blue} Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters for unit tracking. Although they used first-person shooter (FPS) games for their experimentation, the results apply to RTS games as well}. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. \cite{SynnaeveTactics} combined space abstraction into regions from \cite{Perkins10} and tactical-decision making by assigning scores (economical, defenses, etc.) to regions and looking for their correspondences to tactical moves (attacks) in pro-gamers replays.
{\color{blue} Finally, Miles \cite{miles2006co} created the idea of {\em IMTrees}, a tree where each leaf node is an influence map, and each intermediate node is a combination operation (sum, multiplication); Miles used evolutionary algorithms to learn IMTrees for each strategic decision in the game involving spatial reasoning by combining a set of basic influence maps.}

@@

Additionally, there have been some interesting uses of reinforcement learning (RL) \cite{Sutton}: 
Wender and Watson \cite{WenderRL} evaluated the different major RL algorithms for (decentralized) micro-management, which perform all equally. Marthi et al. \cite{Marthi05} employ concurrent hierarchical Q-learning (units Q-functions are combined at the group level) RL to efficiently control units in a ``one robot with multiple effectors'' fashion. Madeira et al. \cite{Madeira06} advocate the use of prior domain knowledge to allow faster RL learning and applied their work on a turn-based strategy game. This is because the action space to explore is gigantic for real game setups. It requires exploiting the existing structure of the game in a partial program (or a partial Markov decision process) and a shape function (or a heuristic) \cite{Marthi05}. Another approach has been proposed by Jaide and Mu{\~n}oz-Avila \cite{jaidee2012classq} through learning just one Q-function for each unit type, in order to cut down the search space. Other approaches that aim at learning the parameters of an underlying model have also been explored. For example Ponsen and Spronck \cite{GA} used evolutionary learning techniques, but face the same problem of dimensionality. For example, evolutionary optimization by simulating fights can easily be adapted to any parameter-dependent micro-management control model, as shown by \cite{OthmanSimu} which optimizes an AIIDE 2010 micro-management competition bot.

@@

Holistic approaches to address RTS AI attempt to address the whole problem using a single unified method. To the best of our knowledge, with a few exceptions, such as the Darmok system \cite{OntanonMSR10} (which uses a combination of case-based reasoning and learning from demonstration) or ALisp \cite{Marthi05}, there has not been much work in this direction.   The main reason is that the complexity of RTS games is too large, and approaches that decompose the problem into smaller, separate, problems, achieve better results in practice. However, holistic approaches, based, for example, on Monte Carlo Tree Search, have only been explored in the context of smaller-scale RTS games~\cite{ontanon2013naive}, but techniques that scale up to large RTS games as StarCraft are still not available.
